Model Recognition as Planning

### 2018-11-18 doc/icaps-19/model-recognition.tex
minor changes by Diego

For each regular language, we generated $20$ random strings, $\mathcal{O}(\tau)$ observations, which length range from 20 to 30 symbols, thus generating a total of 100 strings. More precisely, $\mathcal{O}(\tau)$ are observations in which the applied transitions (actions) are unobservable and the fluents of the automata internal state are also unobservable. Consequently, $\mathcal{O}(\tau)$ are $\Phi$-observations where the problem states are only represented with the fluents that specify the string to classify.

@@

\subsection{Recognizing failures in a non-deterministic {\em blocksworld}}
With this second experiment we aim to validate the two assumptions underlying our approximation of $P(\mathcal{O}|\mathcal{M})$: (1) that agents act rationally and so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation $\mathcal{O}(\tau)$ and (2) that there exists a single such trajectory. Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions. In order to show the assumptions validation:

\begin{itemize}
\item we will measure the {\em  accuracy} of our approximation of $P(\mathcal{O}|\mathcal{M})$ across different degrees of observability; the accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a fluent in an intermediate state of $\mathcal{O}(\tau)$
\item we will assume that the length of the trajectory is known so as to force the optimal explanation of $\mathcal{O}(\tau)$; more precisely, at least one fluent of every state of $\mathcal{O}(\tau)$ is observed and the length of the plan trajectory is thus fixed by the observation
\end{itemize}

@@

$M$ comprises three different {\em extended} 4-schema {\em blocksworld} models. Each model in $M$ includes an additional action schema that encodes one of the three above failures. This means that failures are identified by finding the model that better explains the input observation.

Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations. The increasing trend of the accuracy as observability increases, and so the number of possible explanation lowers, validates our two assumptions.

%The overall high accuracy validates our assumption that the trajectory that better explains the observation is the shortest one. %We limited the failures to be of one kind per observation however, it would be possible to recognize different combinations of them.

>>>>

For each regular language, we generated $20$ random strings, $\mathcal{O}(\tau)$ observations, which lengths range from 20 to 30 symbols, thus generating a total of 100 strings. More precisely, $\mathcal{O}(\tau)$ are observations in which the applied transitions (actions) are unobservable and the fluents of the automata internal state are also unobservable. Consequently, $\mathcal{O}(\tau)$ are $\Phi$-observations where the problem states are only represented with the fluents that specify the string to classify.

@@

\subsection{Recognizing failures in a non-deterministic {\em blocksworld}}
With this second experiment we aim to validate the two assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood: (1) that agents act rationally and so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation $\mathcal{O}(\tau)$ and (2) that there exists a single such trajectory. Our assumptions are more correct with higher observability and become tautological when the trajectory is fully observed ($\mathcal{O(\tau)} = \tau$). In order to show the assumptions validation:

% Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions. In order to show the assumptions validation:

\begin{itemize}
\item we will measure the {\em  accuracy} of our approximation of $P(\mathcal{O}|\mathcal{M})$ across different degrees of observability; the accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a fluent in an intermediate state of $\mathcal{O}(\tau)$
\item we will assume that the length of the trajectory is known so as to keep the possible explanations bounded and force the optimal explanation of $\mathcal{O}(\tau)$; more precisely, at least one fluent of every state of $\mathcal{O}(\tau)$ is observed and the length of the plan trajectory is thus fixed by the observation
\end{itemize}

@@

$M$ comprises three different 4-schema {\em blocksworld} models, each one extended with an additional action schema that encodes one of the three above failures. This means that failures are identified by finding the model that better explains the input observation.

Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations. The positive trend seen in the figure proves that the more complete the observations are the more accurate our estimation becomes.

%The overall high accuracy validates our assumption that the trajectory that better explains the observation is the shortest one.

%The increasing trend of the accuracy as observability increases and so the number of possible explanation lowers, validates our two assumptions.

%We limited the failures to be of one kind per observation however, it would be possible to recognize different combinations of them.
